Guided Proofreading of Automatic Segmentations for Connectomics

### 2016-11-15 paper/guidedproofreading_review.tex
Edited introduction

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
%
Automatic cell image segmentation methods in connectomics produce merge and
split errors, which require correction through proofreading. Previous research
has identified the visual search for these errors as the bottleneck in
interactive proofreading. To aid error correction, we develop two classifiers
that automatically recommend candidate merge and splits to the user. These
classifiers use a convolutional neural network (CNN) that has been trained with
errors in automatic segmentations against expert-labeled ground truth. Our
classifiers detect potentially-erroneous regions by considering a large context
region around a segmentation boundary. Corrections can then be performed by a
user with yes/no decisions  resulting in faster correction times than previous
proofreading methods. We also present a fully-automatic mode that uses a
threshold to make merge/split decisions. Extensive experiments using the
automatic approach and comparing performance of novice and expert users
demonstrate that our method performs favorably agains state-of-the-art
proofreading methods on different connectomics datasets.
%
\end{abstract}

>>>>

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,linkcolor=blue,citecolor=blue,bookmarks=false]{hyperref}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
%
Automatic cell image segmentation methods in connectomics produce merge and
split errors, which require correction through proofreading. Previous research
has identified the visual search for these errors as the bottleneck in
interactive proofreading. To aid error correction, we develop two classifiers
that automatically recommend candidate merge and splits to the user. These
classifiers use a convolutional neural network (CNN) that has been trained with
errors in automatic segmentations against expert-labeled ground truth. Our
classifiers detect potentially-erroneous regions by considering a large context
region around a segmentation boundary. Corrections can then be performed by a
user with yes/no decisions  resulting in faster correction times than previous
proofreading methods. We also present a fully-automatic mode that uses a
probability threshold to make merge/split decisions. Extensive experiments using
the automatic approach and comparing performance of novice and expert users
demonstrate that our method performs favorably agains state-of-the-art
proofreading methods on different connectomics datasets.
%
\end{abstract}

### 2016-11-15 paper/introduction.tex
Edited introduction

In connectomics, neuroscientists annotate neurons and their connectivity within 3D volumes to gain insight into the functional structure of the brain. Rapid progress in automatic sample preparation and electron microscopy (EM) acquisition techniques has made it possible to image large volumes of brain tissue at nanometer resolution. With a voxel size of $4\times4\times40~\text{nm}^3$, a $1~\text{mm}^3$ volume is 1 petabyte of data. With so much data, manual annotation is not feasible, and automatic annotation methods are needed~\cite{jain2010,Liu2014,GALA2014,kaynig2015large}.

Automatic annotation by segmentation and classification of brain tissue is challenging~\cite{isbi_challenge} and all available methods make errors. This leads to the results being \emph{proofread} by humans. This crucial task serves two purposes: 1) to correct errors in the segmentation, and 2) to increase the body of labeled data from which to train better automatic segmentation methods. Recent proofreading tools provide intuitive user interfaces to browse segmentation data in 2D and 3D and to identify and manually correct errors ~\cite{markus_proofreading,raveler,mojo2,haehn_dojo_2014}. Many kinds of errors exist, such as inaccurate boundaries, but the most common are \emph{split errors}, where a single segment is labeled as two, and \emph{merge errors}, where two segments are labeled as one (Fig.~\ref{fig:merge_and_slit_errors}). With user interaction, split errors can be joined, and the missing boundary in a merge error can be defined with manually-seeded watersheds~\cite{haehn_dojo_2014}. However, even with semi-automatic correction tools, the visual inspection to find errors takes the majority of the time~\cite{proofreading_bottleneck}.

@@

Our goal is to automatically detect split and merge errors to reduce visual inspection time. Further, we wish to propose corrections to the user to accept or reject, to reduce correction time. We name this process \textit{guided proofreading}.

First, we learn a classifier for split errors with a convolutional neural network (CNN). This takes as input patches of membrane segmentation probabilities, cell segmentation masks, and boundary masks, and outputs a probabiltiy score. As we must process large data, this classifier only operates on cell boundaries, which reduces computation over methods which analyze every pixel. For merge errors, we invert and reuse the split classification network, instead asking it to rate a set of generated candidate boundaries which hypothesize a split. We compute corrections for both types of errors.

Second, we propose a greedy algorithm for guided proofreading. Possible erroneous regions are sorted by their score, and a correction is generated for each region. Then, a user works through this list of regions and corrections. In a forced choice setting, the user either selects a correction or skips it to advance to the next region. In an automatic setting, errors with a high probability can be automatically corrected first, given an appropriate probability threshold, after which the user would take over. Finally, to test the limits of performance, we create an oracle that only accepts corrections improving the segmentation based on knowledge of the ground truth. This equals perfect proofreading.

Third, we evaluate these methods on multiple connectomics datasets. For the forced choice setting, we perform a quantitative user study with 20 novice users who have no previous experience of proofreading EM data. We ask participants to proofread a small segmentation volume in a fixed time frame. In a between-subjects design, we compare guided proofreading to the semi-automatic \textit{focused proofreading} approach by Plaza~\cite{focused_proofreading}. In addition, we compare against the manual interactive proofreading tool \textit{Dojo} by Haehn~\etal~\cite{haehn_dojo_2014}. We also asked four domain experts to use guided proofreading and focused proofreading for comparison.

We state our contributions:
\begin{itemize}
\item A CNN-based boundary classifier for split errors, plus a merge error classifier which inverts the split error classifier. This is used to propose merge error corrections, removing the need to manually draw the missing edge. These classifiers perform well without much training data, which is expensive to collect for connectomics data.
\item A guided proofreading approach to correcting segmentation volumes, and an assessment scenario comparing forced-choice interaction with automatic and oracle proofreading.
\item A quantitative user study assessing guided proofreading, which shows that our method is able to reduce segmentation error faster than state-of-the-art semi-automatic tools, across both novice and expert users.
\end{itemize}
Our method applies to all existing automatic segmentation methods which produce a label map. As such, we believe that guided proofreading is a promising direction to proofread segmentations more efficiently to help better tackle large volumes of connectomics imagery.

>>>>

In connectomics, neuroscientists annotate neurons and their connectivity within
3D volumes to gain insight into the functional structure of the brain. Rapid
progress in automatic sample preparation and electron microscopy (EM)
acquisition techniques has made it possible to image large volumes of brain
tissue at nanometer resolution. With a voxel size of
$4\times4\times40~\text{nm}^3$, a cubic millimeter volume is one petabyte of
data. With so much data, manual annotation is not feasible, and automatic
annotation methods are needed~\cite{jain2010,Liu2014,GALA2014,kaynig2015large}.

Automatic annotation by segmentation and classification of brain tissue is
challenging~\cite{isbi_challenge} and all available methods make errors. This
requires that the results are being \emph{proofread} by humans. This crucial
task serves two purposes: 1) to correct errors in the segmentation, and 2) to
increase the body of labeled data from which to train better automatic
segmentation methods. Recent proofreading tools provide intuitive user
interfaces to browse segmentation data in 2D and 3D and to identify and manually
correct errors~\cite{markus_proofreading,raveler,mojo2,haehn_dojo_2014}. Many
kinds of errors exist, such as inaccurate boundaries, but the most common are
\emph{split errors}, where a single segment is labeled as two, and \emph{merge
errors}, where two segments are labeled as one
(Fig.~\ref{fig:merge_and_slit_errors}). With user interaction, split errors can
be joined, and the missing boundary in a merge error can be defined with
manually-seeded watersheds~\cite{haehn_dojo_2014}. However, the visual
inspection to find errors takes the majority of the time, even with
semi-automatic correction tools~\cite{proofreading_bottleneck}.

@@

Our goal is to automatically detect split and merge errors to reduce visual
inspection time. Further, to reduce correction time we wish to propose
corrections to the user to accept or reject. We call this process \textit{guided
proofreading}.

We learn a classifier for split errors with a convolutional neural network
(CNN). This takes as input patches of membrane segmentation probabilities, cell
segmentation masks, and boundary masks, and outputs a probability score. As we
must process large data, this classifier only operates on cell boundaries, which
reduces computation over methods that analyze every pixel. For merge errors, we
invert and reuse the split classification network, instead asking it to rate a
set of generated candidate boundaries that hypothesize a split. We compute
corrections for both types of errors.

We also propose a greedy algorithm for guided proofreading. Possible erroneous
regions are sorted by their score, and a correction is generated for each
region. Then, a user works through this list of regions and corrections. In a
forced choice setting, the user either selects a correction or skips it to
advance to the next region. In an automatic setting, errors with a high
probability can be automatically corrected first, given an appropriate
probability threshold, after which the user would take over. Finally, to test
the limits of performance, we create an oracle that only accepts corrections
improving the segmentation based on knowledge of the ground truth. This equals
perfect proofreading.

We evaluate these methods on multiple connectomics datasets. For the forced
choice setting, we perform a quantitative user study with 20 novice users who
have no previous experience of proofreading EM data. We ask participants to
proofread a small segmentation volume in a fixed time frame. In a
between-subjects design, we compare guided proofreading to the semi-automatic
\textit{focused proofreading} approach by Plaza~\cite{focused_proofreading}. In
addition, we compare against the manual interactive proofreading tool
\textit{Dojo} by Haehn~\etal~\cite{haehn_dojo_2014}. We also asked four domain
experts to use guided proofreading and focused proofreading for comparison.

In this paper, we make the following contributions.
%
First, we present a CNN-based boundary classifier for split errors, plus a merge
error classifier that inverts the split error classifier. This is used to
propose merge error corrections, removing the need to manually draw the missing
edge. These classifiers perform well without much training data, which is
expensive to collect for connectomics data.
%
Second, we developed a guided proofreading approach to correcting segmentation
volumes, and an assessment scenario comparing forced-choice interaction with
automatic and oracle proofreading.
%
Furthermore, we present the results of a quantitative user study assessing
guided proofreading, which shows that our method is able to reduce segmentation
error faster than state-of-the-art semi-automatic tools for both novice and
expert users.
%
Our method is applicable to all existing automatic segmentation methods that
produce a label map. As such, we believe that guided proofreading is a promising
direction to proofread segmentations more efficiently to better tackle large
volumes of connectomics imagery.
